---
title: "R for Beginners - R Code File -2"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
urlcolor: blue
---




This R code book has been written by [Rohit Dhankar](https://github.com/RohitDhankar) . GitHub - https://github.com/RohitDhankar 


This is the 2nd in series of R Code Files.     


Refer GitHub Repository , for all Data Files --> https://github.com/RohitDhankar/R-Beginners-Online-Virtual-Learning-Session    


Its a good practice from time to time to keep a track of our current Working Directory and list out all the Objects in our R ENVIRONMENT - specially so when we are committing changes to a Git Remote.   

# VECTOR Operations

```{r}

getwd()
#
ls()

```

We could remove any object with command - rm("Object Name")  

We can also use print() , to view any objects stored value.  


```{r}
# Code Section -1 
a1 <- "FINANCE"
b1 <- "MARKETING"
c1 <- "SALES"
d1 <- 3.1416
char_vector <- c("x","d","c","f")
print(a1)
#
print(char_vector)


```

Going further with VECTORS .   

We combine two or more vectors to get another vector .  


```{r}
# Code Section -2
num_vector <- c(22,22,33,33,44)
print(num_vector)
num_vector1 <- c(11,12,13,14,15)
#
num_vector3 <- c(num_vector,num_vector1)
print(num_vector3)
#
sort(num_vector3)
# 
order(num_vector3) # Ascending Order of  ELEMENTS without SORTING .
#
# The COLON Operator is same as the seq() function seen later in this text. 

seq_1<- 55:50
seq_1
#
seq_2<- 50:55
seq_2
#
# While the above - seq_1 and seq_2 are stored as Objects withing persistence storage. 

50:55 ## is in Memory Only and Not Stored on any persistence storage. 

## Source -- R Manual -- https://stat.ethz.ch/R-manual/R-devel/library/base/html/Colon.html

```

Some basic Maths and Stats with VECTORS.  


```{r}
# Code Section -3
num_vector3 + 5
# Adds NUMERIC VALUE = 5 to all ELEMENTS of the Num Vector. 
nmv_1<-c(20,21,211,312,413,5114)
nmv_2<-c(20,21,211,313,414,5214)
#
class(nmv_1) # "numeric" Vector 
#
print(1/nmv_1)
#
min_max_nmv <- c(min(nmv_1),max(nmv_1))
min_max_nmv  ## Output - MIN == 20 , MAX ==5114
#

### MATHEMATICAL logical operators and Boolean calculus - present in R . 
# - <, <=, >, >=, == for exact equality and != for inequality. 

### Boolean calculus
# Given - nmv_1 and nmv_2 are logical expressions, 
# thus nmv_1 & nmv_2 is intersection ("AND")
# nmv_1 | nmv_2 is union ("OR")
# !nmv_1 is Negation of nmv_1.

nmv_1 == nmv_2

#

nmv_1 != nmv_2

# 
# Lets introduce NA's - the data wranglers nightmare
#
nmv_3 <- c(11,NA,22,33,44,NA,NA)
nmv_3
# We can use function is.na() , to find out the NOT AVAILABLE missing values 
# At a letr stage we shall also look at NA management or IMPUTATION of MISSING VALUES 
# Here is a prelim resource -- 
#
is.na(nmv_3)
#
# Kaggle_Titanic [Multiple Imputation of Missing_Values] -- 
# http://datasciencewithrandpython.blogspot.in/2017/01/kaggle-titanic-initial-analysis-wip.html
#
# Not a NUMBER = NaN
xx <- 0/0.00
xx
#
is.nan(xx) ## TRUE
#
is.nan(nmv_3) ## As Many FALSE - as Elements 
#
nmv_4 <- nmv_3 + 2 ## Any OPERATION done with a NA value , results in a NA value. 
nmv_4
#
```



```{r}
# Code Section -4

num_vector1 * num_vector3
# First 5 elements of - num_vector3 multiplied by the Five Elements 
# of num_vector1 and again the Next 5 elements of num_vector3 
# multiplied by the Five Elements of num_vector1

```

Check out the LENGTH of a VECTOR with length()   


```{r}
# Code Section -5

length(num_vector1 * num_vector3)
```




```{r}
# Code Section -6

#num_vector1 %*% num_vector3 # Error in num_vector1 %*% num_vector3 : non-conformable arguments

# Vectors are not of same Length above - below they are of same length 

nv <- c(1,2,3,4,5)
nv1 <- c(6,7,8,9,10)

nv %*% nv1 # Inner Product of same Length Vectors

# Algeberic Dot Product as defined by WikiPedia - "https://en.wikipedia.org/wiki/Dot_product"


```

Operate upon a ELEMENT of the Vector.  

```{r}
# Code Section -7


log(num_vector3[2]) # Log Base 2 of 22 
#
log(22)
#


```

Converting a CHAR Vector into a NUMERIC Vector .   


```{r}
# Code Section -8
ch_v <- c("11","12","13","14","15")
#
class(ch_v)
#ch_v + 2 # Error in ch_v + 2 : non-numeric argument to binary operator
# Cant do a Math operation on CHAR Vector - lets Convert into NUM Vector 
#
nm_v <- as.numeric(ch_v)
#
class(nm_v)
nm_v + 2 
#
#Summary of the Num Vector as below :- 
#
summary(nm_v+2)
#
summary(nm_v+5)
#
sum(nm_v+5)
#
sd(nm_v+5)
#
max(nm_v+5)
#
min(nm_v+5)
#
mean(nm_v+5)
#
median(nm_v+5)
#
#The Quantile - 
#
quantile(nm_v+5)
#
quantile(nm_v+100)

#
#We can also specify the Quantile buckets or Percentiles as an argument to the Quantile function :-
#
nmv_q <- c(10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90,100)
percent_1 <- quantile(nmv_q, c(.50,.75,.84, .97, .99))
percent_1 
boxplot(percent_1,col = "red",horizontal = TRUE,
        main = "Box and Whisker Plot of Quantiles",
        xlab = "Quantile Values")
# Kindly note how the ARGUMENTs to boxplot()
# have been bumped to the next row - keeping in mind 
# the Horizontal space of our PDF knit of the .Rmd file 

# Seen above we have the MEDIAN quartile - 50% and the UPPER 
# Quartile - 75% along with THREE more percentiles. 
```

Wiki reference -- Percentile Rank -  "https://en.wikipedia.org/wiki/Percentile_rank" 
#

# Intro to ANOVA and BOXPLOTS 

We also carry out ONE Way ANOVA or ANALYSIS of VARIANCE test with the BOX 
and WHISKERS plots as seen below :- 

```{r}
# Code Section -9
library(graphics)

nmv_q <- c(10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90,100)
percent_1 <- quantile(nmv_q, c(.50,.75,.84, .97, .99))
percent_1 

percent_2 <- quantile(nmv_q, c(.1, .3, .16, .40, .50))
percent_2 

percent_3 <- quantile(nmv_q, c(.16, .40, .50,.75,.84))
percent_3 

col_boxes = (c("red","blue","green"))

boxplot(percent_1,percent_2,percent_3,col = col_boxes,
        names = c("perc_1","perc_2","perc_3"),horizontal = TRUE,
        main = "Box and Whisker Plot of Quantiles",
        xlab = "Quantile Values")

# Kindly note the Quantiles are randomly chosen here 
# this is not the best way to choose quantiles 
# we shall come back for details later in this text
```

# rainbow() for Coloring Boxplots - 

```{r}

# Code Section -10

percent_4 <- quantile(nmv_q, c(.16, .40, .50,.95,.99))
percent_4 


percent_5 <- quantile(nmv_q, c(.16, .24,.32 ,.40,.75))
percent_5 

percent_6 <- quantile(nmv_q, c(.1, .5, .26, .45, .60))
percent_6 


percent_7 <- quantile(nmv_q, c(.3, .7, .18, .43, .70))
percent_7 


col_rainbow <- rainbow(7)

boxplot(percent_1,percent_2,percent_3,percent_4,percent_5,percent_6,percent_7,col = col_rainbow,
        names = c("perc_1","perc_2","perc_3","perc_4","perc_5","perc_6","perc_7"),horizontal = TRUE,
        main = "Box and Whisker Plot of Quantiles",
        xlab = "Quantile Values")


```




```{r}
# Code Section -11

# Just for Fun a PIE Graph --- you always 
# need to avoid PIE Graphs 


pie(rep(1, 7), col = rainbow(7))

```

# MATRICE Operations - TRANSPOSE of a MATRIX 

Coming back to MATRICES lets further look at some MATRIX Operations :- 


```{r}
# Code Section -12

m1 <- matrix(data=66:69,nrow=2,ncol=2)
m1

# Lets now TRANSPOSE this MATRIX - for more on TRANSPOSE of MATRICES 
# kindly refer this Wiki Link -- https://en.wikipedia.org/wiki/Transpose


t(m1)

# As seen below - the DIAGONAL Elements remain as -is . 
# 66 and 69 do not move . 
# 67 and 68 switch places , thus giving us a Transpose Matrix. 

# Lets look at another example of TRANSPOSE ....

m2 <- matrix(data=10:25,nrow=4,ncol=4) 
m2
class(m2)

## Note  in the above sequence - 10:25 - both 10 and 25 are included. 
# Lets now TRANSPOSE this MATRIX - for more on TRANSPOSE of MATRICES 
# kindly refer this Wiki Link -- https://en.wikipedia.org/wiki/Transpose

t(m2)

# As seen below - the DIAGONAL Elements remain as-is. 
# 10, 15 , 20 , 25 -- do not move . 
# Non Diagonal elements are Transposed ,giving the Transpose Matrix. 

```
# The Semicolon Notation - RANGE  or SEQUENCE

```{r}
# Code Section -13

# Quick recap of the SEQUENCE 

a_seq <- 66:69 
a_seq 

# In the earlier section we create a MATRIX by using a sequence within 
# the COMBINE function

# We can also use the - seq - sequence function as seen below 

b_seq <- seq(from=66, to=69, by=1)
b_seq

#

b_seq <- seq(from=66, to=69, by=2)
b_seq

#

c_seq <- seq(from=1, to=10, by=2)
c_seq
class(c_seq)

```

# The CBIND and RBIND Functions 

We can COLUMN Bind and ROW Bind more than one data structures as seen below - 
```{r}

# Code Section -14

ma1 <- matrix(data=10:15,nrow=3,ncol=2) 
ma1
#
class(ma1)
#
ma2 <- matrix(data=20:25,nrow=3,ncol=2) 
ma2
#
class(ma2)

# ROW Bind the Matrices 
ma3 <- rbind(ma1,ma2)
ma3
#
# COLUMN Bind the Matrices 
ma4 <- cbind(ma1,ma2)
ma4

# As seen below we need to have same COLUMN Numbers to do a RBIND
#m3 <- rbind(m1,m2)

# # As seen below we need to have same ROW Numbers to do a RBIND
#m3 <- cbind(m1,m2)

```

# ROW BIND - Data Frames

```{r}

# Code Section -15

df_1 <- read.csv("~/Desktop/R_Own/R_1 - Sheet1.csv",header =TRUE , sep = "," )
df_1

```


```{r}
# Code Section -16

df_2 <- read.csv("~/Desktop/R_Own/R_2.csv",header =TRUE , sep = "," )
df_2


```
# COLUMN BIND- Data Frame

```{r}

# Code Section -17

df_3_C <- cbind(df_1,df_2)
df_3_C

```

# ROW BIND - Data Frame 

```{r}

# Code Section -18

df_3_R <- rbind(df_2,df_1)
df_3_R

```


# describeBy() with Grouping variables 

```{r}
# Code Section -19

library("psych", lib.loc="~/R/x86_64-pc-linux-gnu-library/3.3")
library(psych)

#?describeBy() # Seek HELP for the function - Uncomment this line. 

describeBy(df_3_C,df_3_C$Date.of.Invoice) 

# In this case - Grouping Variable is - Date of Invoice . 
# This Grouping variable has two Values here - 01-01-2013 and 02-01-2013

# IMPORTANT NOTE --- The DOLLAR SIGN in df_3_C$Date.of.Invoice , is used 
# to access a certain variable within the DATA FRAME. 

# Also note that the variables within the DATA FRAME will not be stored with the Labels 
# as - is 

```

# Built in DATA Sets 

The default R environment - comes bundled with a number of packages and data sets. 

A package called DATASETS - contains a number of inbuilt data sets as seen in sections below - 

```{r}
# Code Section -20

library(help = "datasets")


# 
# Seen below are the DataSets available within my installed version of R - 
# yours may slightly differ :- 
# 
# AirPassengers           Monthly Airline Passenger Numbers 1949-1960
# BJsales                 Sales Data with Leading Indicator
# BOD                     Biochemical Oxygen Demand
# CO2                     Carbon Dioxide Uptake in Grass Plants
# ChickWeight             Weight versus age of chicks on different diets
# DNase                   Elisa assay of DNase
# EuStockMarkets          Daily Closing Prices of Major European Stock
#                         Indices, 1991-1998
# Formaldehyde            Determination of Formaldehyde
# HairEyeColor            Hair and Eye Color of Statistics Students
# Harman23.cor            Harman Example 2.3
# Harman74.cor            Harman Example 7.4
# Indometh                Pharmacokinetics of Indomethacin
# InsectSprays            Effectiveness of Insect Sprays
# JohnsonJohnson          Quarterly Earnings per Johnson & Johnson Share
# LakeHuron               Level of Lake Huron 1875-1972
# LifeCycleSavings        Intercountry Life-Cycle Savings Data
# Loblolly                Growth of Loblolly pine trees
# Nile                    Flow of the River Nile
# Orange                  Growth of Orange Trees
# OrchardSprays           Potency of Orchard Sprays
# PlantGrowth             Results from an Experiment on Plant Growth
# Puromycin               Reaction Velocity of an Enzymatic Reaction
# Theoph                  Pharmacokinetics of Theophylline
# Titanic                 Survival of passengers on the Titanic
# ToothGrowth             The Effect of Vitamin C on Tooth Growth in
#                         Guinea Pigs
# UCBAdmissions           Student Admissions at UC Berkeley
# UKDriverDeaths          Road Casualties in Great Britain 1969-84
# UKLungDeaths            Monthly Deaths from Lung Diseases in the UK
# UKgas                   UK Quarterly Gas Consumption
# USAccDeaths             Accidental Deaths in the US 1973-1978
# USArrests               Violent Crime Rates by US State
# USJudgeRatings          Lawyers' Ratings of State Judges in the US
#                         Superior Court
# USPersonalExpenditure   Personal Expenditure Data
# VADeaths                Death Rates in Virginia (1940)
# WWWusage                Internet Usage per Minute
# WorldPhones             The World's Telephones
# ability.cov             Ability and Intelligence Tests
# airmiles                Passenger Miles on Commercial US Airlines,
#                         1937-1960
# airquality              New York Air Quality Measurements
# anscombe                Anscombe's Quartet of 'Identical' Simple Linear
#                         Regressions
# attenu                  The Joyner-Boore Attenuation Data
# attitude                The Chatterjee-Price Attitude Data
# austres                 Quarterly Time Series of the Number of
#                         Australian Residents
# beavers                 Body Temperature Series of Two Beavers
# cars                    Speed and Stopping Distances of Cars
# chickwts                Chicken Weights by Feed Type
# co2                     Mauna Loa Atmospheric CO2 Concentration
# crimtab                 Student's 3000 Criminals Data
# datasets-package        The R Datasets Package
# discoveries             Yearly Numbers of Important Discoveries
# esoph                   Smoking, Alcohol and (O)esophageal Cancer
# euro                    Conversion Rates of Euro Currencies
# eurodist                Distances Between European Cities and Between
#                         US Cities
# faithful                Old Faithful Geyser Data
# freeny                  Freeny's Revenue Data
# infert                  Infertility after Spontaneous and Induced
#                         Abortion
# iris                    Edgar Anderson's Iris Data
# islands                 Areas of the World's Major Landmasses
# lh                      Luteinizing Hormone in Blood Samples
# longley                 Longley's Economic Regression Data
# lynx                    Annual Canadian Lynx trappings 1821-1934
# morley                  Michelson Speed of Light Data
# mtcars                  Motor Trend Car Road Tests
# nhtemp                  Average Yearly Temperatures in New Haven
# nottem                  Average Monthly Temperatures at Nottingham,
#                         1920-1939
# npk                     Classical N, P, K Factorial Experiment
# occupationalStatus      Occupational Status of Fathers and their Sons
# precip                  Annual Precipitation in US Cities
# presidents              Quarterly Approval Ratings of US Presidents
# pressure                Vapor Pressure of Mercury as a Function of
#                         Temperature
# quakes                  Locations of Earthquakes off Fiji
# randu                   Random Numbers from Congruential Generator
#                         RANDU
# rivers                  Lengths of Major North American Rivers
# rock                    Measurements on Petroleum Rock Samples
# sleep                   Student's Sleep Data
# stackloss               Brownlee's Stack Loss Plant Data
# state                   US State Facts and Figures
# sunspot.month           Monthly Sunspot Data, from 1749 to "Present"
# sunspot.year            Yearly Sunspot Data, 1700-1988
# sunspots                Monthly Sunspot Numbers, 1749-1983
# swiss                   Swiss Fertility and Socioeconomic Indicators
#                         (1888) Data
# treering                Yearly Treering Data, -6000-1979
# trees                   Girth, Height and Volume for Black Cherry Trees
# uspop                   Populations Recorded by the US Census
# volcano                 Topographic Information on Auckland's Maunga
#                         Whau Volcano
# warpbreaks              The Number of Breaks in Yarn during Weaving
# women                   Average Heights and Weights for American Women
# 


```

# ATTACH and DETACH Datasets 

To OPERATE upon an inbuilt DATASET we need to LOAD it into R . 

In lay-man terms , we may equate this to our earlier - read.csv() , but 

there are subtle differences. 

The ATTACH will pull in a Data Set from the DataSet Package into our R Search Path. 


```{r}
# Code Section -21

?attach()

# # Quoting below from inbuilt HELP - 
# # The database is attached to the R search path. This means that the database is 
# searched by R when evaluating a variable, so objects in the database can be accessed 
# by simply giving their names.

?detach()

# # Quoting below from inbuilt HELP - 
# # Detach a database, i.e., remove it from the search() path of available R objects. 
# Usually this is either a data.frame which has been attached or a 
# package which was attached by library.

# Another Function which will help automate is - zap()

# library(epicalc)
# ?zap()
# https://artax.karlin.mff.cuni.cz/r-help/library/epicalc/html/zap.html
```



```{r}
# Code Section -22

search()

```

# DATA_FRAME COLUMNS - The DF Column Vector


```{r}
# Code Section -23

# Three basic methods to access a DF COLUMN - also called the 
# DF Column Vector

print(df_1[[2]]) # Will Print 2nd COLUMN of the DF 
#
print(df_1[["Product.Name"]]) # Will Print the "NAMED"  COLUMN of the DF 
#
print(df_1$Product.Name) # Will Print the "NAMED" COLUMN of the DF 


```

# DATA_FRAME - Row and Column Slices

```{r}
# Code Section -24

# Slicing the DF COLUMNS and ROWS 

print(df_1[2]) # Will Print 2nd COLUMN - ALL ROWS - of the DF as a SLICE. 
#
print(df_1[2,]) # Will Print 2nd ROW - ALL COLUMNS - of the DF as a SLICE. 
#
print(df_1[c(2,4,6)]) # Using an INDEX VECTOR  - created with COMBINE Function. 
# To access MULTIPLE COLUMNS - 2,4 and 6 of the DF.

```

Lets create a New Directory - Files Folder from within R and house our 
basic introductory graphs there - we use - dir.create()


```{r}
# Code Section -25

#dir.create("/home/dhankar/Desktop/R_Own/Plots/Intro_1/", recursive=TRUE) # to be RUN only Once

```

# Plots and Graphs - the Basics 

Over the next couple of sections we shall preview the basics of Plots and Graphs. 

If we were to execute the code from the following sections , in the Console- the
Graphs would not appear inline but would appear within the PLOTS pane 
of R Studio or in certain cases as  POP Up's . 


# HISTOGRAM an Introduction 


```{r}
# Code Section -26
# So what exactly is a HISTOGRAM - 
# " estimate of the probability distribution of a continuous variable (quantitative variable)"
# Source -- Wiki -- https://en.wikipedia.org/wiki/Histogram


hist(df_1$Cost.Price,col.main="blue",col = "green")
#hist(df_1$Cost.Price,df_1$Sales.Price)
#

hist(df_1$Sales.Price,col.main="blue",col = "orange")

```



```{r}
# Code Section -27



```



```{r}
# Code Section -28


```




```{r}
# Code Section -29


```





```{r}
# Code Section -30


```




```{r}
# Code Section -32


```







```{r}
# Code Section -33


```






```{r}
# Code Section -34


```








```{r}
# Code Section -35


```








```{r}
# Code Section -36


```





```{r}
sessionInfo()

```



