---
title: "R_Code_1.Rmd"
author: "Rohit Dhankar"
date: "20 July 2017"
output:
  pdf_document: default
  html_document: default
---

This is the 2nd in series of R Code Files.     


Refer the GitHub Repository , for all Code files ---> https://github.com/RohitDhankar/R-Beginners-Online-Virtual-Learning-Session    


Its a good practise from time to time to keep a track of our current Working Directory and list out all the Objects in our R ENVIRONMENT - specially so when we are committing changes to a Git Remote.   


```{r}

getwd()
#
ls()

```

We could remove any object with command - rm("Object Name")  

We can also use print() , to view any objects stored value.  


```{r}
# Code Section -1 
#print(a1)
#
#print(char_vector)
#

```

Going further with VECTORS .   

We combine two or more vectors to get another vector .  


```{r}
# Code Section -2
num_vector <- c(22,22,33,33,44)
print(num_vector)
num_vector1 <- c(11,12,13,14,15)
#
num_vector3 <- c(num_vector,num_vector1)
print(num_vector3)
```

Some basic Maths and Stats with VECTORS.  


```{r}
# Code Section -3
num_vector3 + 5

# Adds NUMERIC VALUE = 5 to all ELEMENTS of the Num Vector. 

```



```{r}
# Code Section -4

num_vector1 * num_vector3
# First 5 elements of - num_vector3 multiplied by the Five Elements of num_vector1 and again the Next 5 elements of num_vector3 multiplied by the Five Elements of num_vector1

```

Check out the LENGTH of a VECTOR with length()   


```{r}
# Code Section -5

length(num_vector1 * num_vector3)
```




```{r}
# Code Section -6

#num_vector1 %*% num_vector3 # Error in num_vector1 %*% num_vector3 : non-conformable arguments

# Vectors are not of same Length above - below they are of same length 

nv <- c(1,2,3,4,5)
nv1 <- c(6,7,8,9,10)

nv %*% nv1 # Inner Product of same Length Vectors

# Algeberic Dot Product as defined by WikiPedia - "https://en.wikipedia.org/wiki/Dot_product"


```

Operate upon a ELEMENT of the Vector.  

```{r}
# Code Section -7


log(num_vector3[2]) # Log Base 2 of 22 
#
log(22)
#


```

Converting a CHAR Vector into a NUMERIC Vector .   


```{r}
# Code Section -8
ch_v <- c("11","12","13","14","15")
#
class(ch_v)
#ch_v + 2 # Error in ch_v + 2 : non-numeric argument to binary operator
# Cant do a Math operation on CHAR Vector - lets Convert into NUM Vector 
#
nm_v <- as.numeric(ch_v)
#
class(nm_v)
nm_v + 2 
#
print("Summary of the Num Vector as below :- ")
summary(nm_v+2)
#
summary(nm_v+5)
#
sum(nm_v+5)
#
sd(nm_v+5)
#
max(nm_v+5)
#
min(nm_v+5)
#
mean(nm_v+5)
#
median(nm_v+5)
#

```



```{r}
# Code Section -9

plot(nm_v + 5)

```




```{r}
# Code Section -10


```




```{r}
# Code Section -11

```



```{r}
# Code Section -12

```



```{r}
# Code Section -13

```



```{r}
# Code Section -14

```



```{r}
# Code Section -15

```



```{r}
# Code Section -16

```



```{r}
# Code Section -17

```


sessionInfo()
  
R version 3.3.2 (2016-10-31)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Ubuntu 16.04.1 LTS

locale:
 [1] LC_CTYPE=en_IN.UTF-8       LC_NUMERIC=C               LC_TIME=en_IN.UTF-8        LC_COLLATE=en_IN.UTF-8    
 [5] LC_MONETARY=en_IN.UTF-8    LC_MESSAGES=en_IN.UTF-8    LC_PAPER=en_IN.UTF-8       LC_NAME=C                 
 [9] LC_ADDRESS=C               LC_TELEPHONE=C             LC_MEASUREMENT=en_IN.UTF-8 LC_IDENTIFICATION=C       

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

loaded via a namespace (and not attached):
 [1] backports_1.0.4 magrittr_1.5    rprojroot_1.1   htmltools_0.3.5 tools_3.3.2     base64enc_0.1-3 yaml_2.1.14    
 [8] Rcpp_0.12.8     stringi_1.1.2   rmarkdown_1.3   knitr_1.15.1    jsonlite_1.1    stringr_1.1.0   digest_0.6.10  
[15] evaluate_0.10  

  
EOF - R_Code_1.Rmd

