---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---


This is the 3rd in series of R Code Files.     


Refer the GitHub Repository , for all Code files ---> https://github.com/RohitDhankar/R-Beginners-Online-Virtual-Learning-Session    


Its a good practice from time to time to keep a track of our current Working Directory and list out all the Objects in our R ENVIRONMENT - specially so when we are committing changes to a Git Remote.   

# Heading -1 

```{r}


```


We can also use print() , to view any objects stored value.  


```{r}
# Code Section -1 



```



```{r}
# Code Section -2


```

# 


```{r}
# Code Section -3



```



```{r}
# Code Section -4


```




```{r}
# Code Section -5


```




```{r}
# Code Section -6

#num_vector1 %*% num_vector3 # Error in num_vector1 %*% num_vector3 : non-conformable arguments

# Vectors are not of same Length above - below they are of same length 

nv <- c(1,2,3,4,5)
nv1 <- c(6,7,8,9,10)

nv %*% nv1 # Inner Product of same Length Vectors

# Algeberic Dot Product as defined by WikiPedia - "https://en.wikipedia.org/wiki/Dot_product"


```

Operate upon a ELEMENT of the Vector.  

```{r}
# Code Section -7


log(num_vector3[2]) # Log Base 2 of 22 
#
log(22)
#


```

Converting a CHAR Vector into a NUMERIC Vector .   


```{r}
# Code Section -8
ch_v <- c("11","12","13","14","15")
#
class(ch_v)
#ch_v + 2 # Error in ch_v + 2 : non-numeric argument to binary operator
# Cant do a Math operation on CHAR Vector - lets Convert into NUM Vector 
#
nm_v <- as.numeric(ch_v)
#
class(nm_v)
nm_v + 2 
#
#Summary of the Num Vector as below :- 
#
summary(nm_v+2)
#
summary(nm_v+5)
#
sum(nm_v+5)
#
sd(nm_v+5)
#
max(nm_v+5)
#
min(nm_v+5)
#
mean(nm_v+5)
#
median(nm_v+5)
#
#The Quantile - 
#
quantile(nm_v+5)
#
quantile(nm_v+100)

#
#We can also specify the Quantile buckets or Percentiles as an argument to the Quantile function :-
#
nmv_q <- c(10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90,100)
percent_1 <- quantile(nmv_q, c(.50,.75,.84, .97, .99))
percent_1 
boxplot(percent_1,col = "red",horizontal = TRUE,
        main = "Box and Whisker Plot of Quantiles",
        xlab = "Quantile Values")
# Kindly note how the ARGUMENTs to boxplot()
# have been bumped to the next row - keeping in mind 
# the Horizontal space of our PDF knit of the .Rmd file 

# Seen above we have the MEDIAN quartile - 50% and the UPPER Quartile - 75% along with THREE more percentiles. 
```

Wiki reference -- Percentile Rank -  "https://en.wikipedia.org/wiki/Percentile_rank" 
#

# Intro to ANOVA and BOXPLOTS 

We also carry out ONE Way ANOVA or ANALYSIS of VARIANCE test with the BOX 
and WHISKERS plots as seen below :- 

```{r}
# Code Section -9
library(graphics)

nmv_q <- c(10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90,100)
percent_1 <- quantile(nmv_q, c(.50,.75,.84, .97, .99))
percent_1 

percent_2 <- quantile(nmv_q, c(.1, .3, .16, .40, .50))
percent_2 

percent_3 <- quantile(nmv_q, c(.16, .40, .50,.75,.84))
percent_3 

col_boxes = (c("red","blue","green"))

boxplot(percent_1,percent_2,percent_3,col = col_boxes,
        names = c("perc_1","perc_2","perc_3"),horizontal = TRUE,
        main = "Box and Whisker Plot of Quantiles",
        xlab = "Quantile Values")

# Kindly note the Quantiles are randomly chosen here 
# this is not the best way to choose quantiles 
# we shall come back for details later in this text
```

# rainbow() for Coloring Boxplots - 

```{r}

# Code Section -10

percent_4 <- quantile(nmv_q, c(.16, .40, .50,.95,.99))
percent_4 


percent_5 <- quantile(nmv_q, c(.16, .24,.32 ,.40,.75))
percent_5 

percent_6 <- quantile(nmv_q, c(.1, .5, .26, .45, .60))
percent_6 


percent_7 <- quantile(nmv_q, c(.3, .7, .18, .43, .70))
percent_7 


col_rainbow <- rainbow(7)

boxplot(percent_1,percent_2,percent_3,percent_4,percent_5,percent_6,percent_7,col = col_rainbow,
        names = c("perc_1","perc_2","perc_3","perc_4","perc_5","perc_6","perc_7"),horizontal = TRUE,
        main = "Box and Whisker Plot of Quantiles",
        xlab = "Quantile Values")


```




```{r}
# Code Section -11

# Just for Fun a PIE Graph --- you always 
# need to avoid PIE Graphs 


pie(rep(1, 7), col = rainbow(7))

```

# MATRICE Operations - TRANSPOSE of a MATRIX 

Coming back to MATRICES lets further look at some MATRIX Operations :- 


```{r}
# Code Section -12

m1 <- matrix(data=66:69,nrow=2,ncol=2)
m1

# Lets now TRANSPOSE this MATRIX - for more on TRANSPOSE of MATRICES 
# kindly refer this Wiki Link -- https://en.wikipedia.org/wiki/Transpose


t(m1)

# As seen below - the DIAGONAL Elements remain as -is . 
# 66 and 69 do not move . 
# 67 and 68 switch places , thus giving us a Transpose Matrix. 

# Lets look at another example of TRANSPOSE ....

m2 <- matrix(data=10:25,nrow=4,ncol=4) 
m2
class(m2)

## Note  in the above sequence - 10:25 - both 10 and 25 are included. 
# Lets now TRANSPOSE this MATRIX - for more on TRANSPOSE of MATRICES 
# kindly refer this Wiki Link -- https://en.wikipedia.org/wiki/Transpose

t(m2)

# As seen below - the DIAGONAL Elements remain as-is. 
# 10, 15 , 20 , 25 -- do not move . 
# Non Diagonal elements are Transposed ,giving the Transpose Matrix. 

```
# The Semicolon Notation - RANGE  or SEQUENCE

```{r}
# Code Section -13

# Quick recap of the SEQUENCE 

a_seq <- 66:69 
a_seq 

# In the earlier section we create a MATRIX by using a sequence within the COMBINE function
# We can also use the - seq - sequence function as seen below 

b_seq <- seq(from=66, to=69, by=1)
b_seq

#

b_seq <- seq(from=66, to=69, by=2)
b_seq

#

c_seq <- seq(from=1, to=10, by=2)
c_seq
class(c_seq)

```

# The CBIND and RBIND Functions 

We can COLUMN Bind and ROW Bind more than one data structures as seen below - 
```{r}

# Code Section -14

ma1 <- matrix(data=10:15,nrow=3,ncol=2) 
ma1
#
class(ma1)
#
ma2 <- matrix(data=20:25,nrow=3,ncol=2) 
ma2
#
class(ma2)

# ROW Bind the Matrices 
ma3 <- rbind(ma1,ma2)
ma3
#
# COLUMN Bind the Matrices 
ma4 <- cbind(ma1,ma2)
ma4

# As seen below we need to have same COLUMN Numbers to do a RBIND
#m3 <- rbind(m1,m2)

# # As seen below we need to have same ROW Numbers to do a RBIND
#m3 <- cbind(m1,m2)

```

# ROW BIND - Data Frames

```{r}

# Code Section -15

df_1 <- read.csv("~/Desktop/R_Own/R_1 - Sheet1.csv",header =TRUE , sep = "," )
df_1

```


```{r}
# Code Section -16

df_2 <- read.csv("~/Desktop/R_Own/R_2.csv",header =TRUE , sep = "," )
df_2


```
# COLUMN BIND- Data Frame

```{r}

# Code Section -17

df_3_C <- cbind(df_1,df_2)
df_3_C

```

# ROW BIND - Data Frame 

```{r}

# Code Section -18

df_3_R <- rbind(df_2,df_1)
df_3_R

```

# describeBy() with Grouping variables 

```{r}
# Code Section -19

library("psych", lib.loc="~/R/x86_64-pc-linux-gnu-library/3.3")
library(psych)

#?describeBy() # Seek HELP for the function - Uncomment this line. 

describeBy(df_3_C,df_3_C$Date.of.Invoice) 

# In this case - Grouping Variable is - Date of Invoice . 
# This Grouping variable has two Values here - 01-01-2013 and 02-01-2013
# 

```


```{r}
# Code Section -20


```



```{r}
# Code Section -21


```



```{r}
# Code Section -22


```



```{r}
# Code Section -23


```



```{r}
# Code Section -24


```





```{r}
# Code Section -25


```



```{r}
# Code Section -26


```



```{r}
# Code Section -27


```



```{r}
# Code Section -28


```




```{r}
# Code Section -29


```





```{r}
# Code Section -30


```




```{r}
# Code Section -32


```







```{r}
# Code Section -33


```






```{r}
# Code Section -34


```








```{r}
# Code Section -35


```








```{r}
# Code Section -36


```





```{r}
sessionInfo()

```



