---
title: "R for Marketing - Market Basket Analysis"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
urlcolor: blue
---


This R code book written by [Rohit Dhankar](https://github.com/RohitDhankar) . GitHub - https://github.com/RohitDhankar 

Code and Data > [Rohit Dhankar-GITHUB](https://github.com/RohitDhankar)

The dataset is Copyrighted by - 

Brijs T., Swinnen G., Vanhoof K., and Wets G. (1999), The use of association rules for prod-
uct assortment decisions: a case study, in: Proceedings of the Fifth International Conference on
Knowledge Discovery and Data Mining, San Diego (USA), August 15-18, pp.  254-260.  ISBN:
1-58113-143-7.

# R for Marketing - Market Basket Analysis

```{r}
# Getting the Retail Sales Data -- http://fimi.ua.ac.be/data/retail.pdf
# Data can also be taken from -- http://r-marketing.r-forge.r-project.org
# Also at -- http://www.cis.hut.fi/Opinnot/T-61.5060/2005/t615060-l-2005-09-29-b.pdf


# Set Seed -- ensure reproducible results 

set.seed(123)

require(arules) # Mining Association Rules and Frequent Itemsets
#fileURL <- "http://fimi.ua.ac.be/data/retail.dat.gz"
#download.file(fileURL, destfile="retail.data.gz", method="curl")
# Read the data in basket format
transaction_data = read.transactions("retail.data.gz", format = "basket", sep=" ");
#
# The Structure of the Data
str(transaction_data)
##
# Highlighted for focus --- 
# .. .. ..@ i
# : int [1:908576] 0 1 2 1113 2224 3335 4446 5557 6668 7249 ...
# ##
# .. .. ..@ p
# : int [1:88163] 0 30 33 36 47 51 64 70 73 79 ...
# ##
# .. .. ..@ Dim
# : int [1:2] 16470 88162
#
# The Summary of the Data 
summary(transaction_data)
 
```
# Data Summary 

1. Each ROW or Obs of the data is - 1 - Market Basket.  
2. Each Market Basket has many different purchase items.  
3. There are - 88,162 ROWS.  
4. Each ROW mostly has different no. of purchase items.  



```{r}
# Using Function - readLines() from BASE 
#
transaction_data1<- readLines("~/Desktop/R_Own/R_Marketing/Mkt_DATA_Files/retail.dat")
#


```



```{r}
# Read the HEAD  end of the Data 
#
head(transaction_data1)
#



```


```{r}
# Read Only the 5th MARKET BASKET and see how many ITEMS in it and WHICH ALL Items ? 
#
head(transaction_data1[5])


```


```{r}
# Read TAIL end of the Data 
#
tail(transaction_data1)


```
We can now create a DATA STRUCTURE with LABELS for each,     
Market basket. Each Market Basket is also called a - TRANSACTION.   

To create LABELS or ROW Names we need a NAMES Vector.  
Lets create a ROW_Labels Vector.  


```{r}
# Code for ROW_Labels vector 
#
ROW_Labels <- paste("Market_basket_",1:length(transaction_data1),sep="")
#
typeof(ROW_Labels)
#
class(ROW_Labels)
#
ROW_Labels[1:20] # Prints initial 20 LABELS 
#
tail(ROW_Labels) # Market_basket_88162"
#
str(ROW_Labels)
# As seen above - we have created a CHARACTER Vector with 
# ROW_Lables == Market_basket_1 to Market_basket_88162


```

We now use the Function - strsplit() , to SPLIT the data at WHITESPACE    
into ROWS , thus whenever a WHITESPACE occurs or a MARKET BASKET or ROW ends  
the data is split into the Next Row .  


```{r}

transaction_list <-strsplit(transaction_data1,"")
#
str(transaction_list)
# 
# We Add the ROW_names Labels 
#
names(transaction_list) <- paste("Market_basket_",1:length(transaction_data1),sep="")
#
# Again see the STRUCTURE
#
str(transaction_list) # 
#
tail(transaction_list)
#
#
# We can now clearly see the SPARSNESS of the Data 
# The "" - Blanks occur eeven adjacent to each other 
# seen in - $Market_basket_88159 - thus whatever Order 
# is maintained in the Brick and Mortar RETAIL STORE 
# is not always suggestive of Association 
# 
# Also looking at $Market_basket_88160
 #    [1] "2" "3" "1" "0" " " "4" "2" "6" "7" " "
# we can observe a FLIP FLOP Pattern 
#


```



```{r}

market_basket_rules <- apriori(transaction_data, parameter=list(supp=0.001, conf=0.4))


```

# DATA VISUALIZATION - A RULES VIZ 

We now visualize the Data to make further sense of the Association Rules.  


```{r}
require(arulesViz)

### Errors 




```




```{r}


```




```{r}


```




```{r}


```




```{r}


```




```{r}


```




```{r}


```




```{r}


```




```{r}


```
