---
title: "R for Beginners - R for Marketing Code File-1"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
urlcolor: blue
---


This R code book written by [Rohit Dhankar](https://github.com/RohitDhankar) . GitHub - https://github.com/RohitDhankar 

Code and Data > https://github.com/RohitDhankar/R-Beginners-Online-Virtual-Learning-Session    

Good practice to keep track of current Working Directory , list all Objects in R ENVIRONMENT - specially so when committing changes to Git or any other version control Remote directory. 

# R for Marketing 

```{r}
# Simulating Synthetic Data 

# Set Seed -- ensure reproducible results 

set.seed(123)

# Presume a retail stores chain called - Mkt , having 200 Stores globally
# Each Country has a store within their capital city
# Do consider this code is NOT DRY :) 
# I need to recode this bit keeping in mind the 
# DONT REPEAT YOURSELF rule. 


# Scalar Vector Constant - tweak to change DF Dimensions
aa<-1500

# Dates on which Data gathered 
# we simulate 10 sets of dates 
# when aa == 1500 , we get 150 dates in each set 
# we then combine these sets into a DATES vector 
# assign this DATES vector 

dates_aa <-seq(as.Date("2000/1/1"), by = "day", length.out = aa/10) 
str(dates_aa)
date_temp_1<-dates_aa[aa/10]
date_temp_1


dates_bb <-seq(as.Date(date_temp_1), by = "day", length.out = aa/10)
str(dates_bb) ; date_temp_2<-dates_bb[aa/10] ; date_temp_2

dates_cc <-seq(as.Date(date_temp_2), by = "day", length.out = aa/10)
str(dates_cc) ; date_temp_3<-dates_cc[aa/10] ; date_temp_3

dates_dd <-seq(as.Date(date_temp_3), by = "day", length.out = aa/10)
str(dates_dd) ; date_temp_4<-dates_dd[aa/10] ; date_temp_4

dates_ee <-seq(as.Date(date_temp_4), by = "day", length.out = aa/10)
str(dates_ee) ; date_temp_5<-dates_ee[aa/10] ; date_temp_5

dates_ff <-seq(as.Date(date_temp_5), by = "day", length.out = aa/10)
str(dates_ff) ; date_temp_6<-dates_ff[aa/10] ; date_temp_6

dates_gg <-seq(as.Date(date_temp_6), by = "day", length.out = aa/10)
str(dates_gg) ; date_temp_7<-dates_gg[aa/10] ; date_temp_7

dates_hh <-seq(as.Date(date_temp_7), by = "day", length.out = aa/10)
str(dates_hh) ; date_temp_8<-dates_hh[aa/10] ; date_temp_8

dates_ii <-seq(as.Date(date_temp_8), by = "day", length.out = aa/10)
str(dates_ii) ; date_temp_9<-dates_ii[aa/10] ; date_temp_9

dates_jj <-seq(as.Date(date_temp_9), by = "day", length.out = aa/10)
str(dates_jj) ; date_temp_10<-dates_jj[aa/10] ; date_temp_10

# CHECK --- Could i have done this faster in Python ?? 


# Func - seq(as.Date ...)
# REFER -- https://stat.ethz.ch/R-manual/R-devel/library/base/html/seq.Date.html

# Mkt Stores ID's == ms_ids

ms_cntry1 <- c(rep("IND",aa))
ms_cntry2 <- c(rep("AUS",aa))
ms_cntry3 <- c(rep("NZ",aa))
ms_cntry4 <- c(rep("RUS",aa))
ms_cntry5 <- c(rep("USA",aa))
ms_cntry6 <- c(rep("MEX",aa))
ms_cntry7 <- c(rep("CAN",aa))
ms_cntry8 <- c(rep("BRZ",aa))
ms_cntry9 <- c(rep("SPN",aa))
ms_cntry10 <- c(rep("FRA",aa))
#
# #
ms_cty1 <- c(rep("CTY_1",aa))
ms_cty2 <- c(rep("CTY_2",aa))
ms_cty3 <- c(rep("CTY_3",aa))
ms_cty4 <- c(rep("CTY_4",aa))
ms_cty5 <- c(rep("CTY_5",aa))
ms_cty6 <- c(rep("CTY_6",aa))
ms_cty7 <- c(rep("CTY_7",aa))
ms_cty8 <- c(rep("CTY_8",aa))
ms_cty9 <- c(rep("CTY_9",aa))
ms_cty10 <- c(rep("CTY_10",aa))
# #
# #
# # Using - runif() # runif generates random deviates.
psale_1 <- runif(aa,min=100,max=120) ## How many values Required the - N == aa
psale_2 <- runif(aa,min=15,max=20) ##
psale_3 <- runif(aa,min=25,max=30) ##
psale_4 <- runif(aa,min=100,max=320) ##
psale_5 <- runif(aa,min=5,max=140) ##
psale_6 <- runif(aa,min=25,max=350) ##
psale_7 <- runif(aa,min=100,max=620) ##
psale_8 <- runif(aa,min=5,max=80) ##
psale_9 <- runif(aa,min=25,max=90) ##
psale_10 <- runif(aa,min=100,max=620) ##
#
# #
# # Using - runif() # runif generates random deviates.
pcost_1 <- runif(aa,min=111.49,max=120.56) ## How many values Required the - N == 5
pcost_2 <- runif(aa,min=65.05,max=100.42) ## Random MINIMUM Value  == 65.05
pcost_3 <- runif(aa,min=500.44,max=3000.78) ## Random MAXIMUM Value == 3000.78
pcost_4 <- runif(aa,min=300.44,max=3000.78) ##
pcost_5 <- runif(aa,min=400.44,max=3000.78) ##
pcost_6 <- runif(aa,min=900.44,max=3000.78) ##
pcost_7 <- runif(aa,min=1100.44,max=37000.78) ##
pcost_8 <- runif(aa,min=1400.44,max=32000.78) ##
pcost_9 <- runif(aa,min=1700.44,max=33000.78) ##
pcost_10 <- runif(aa,min=5500.44,max=30000.78) ##
#
# Data Frame from NUMERIC and CHARACTER VECTORS
#
# p_sale_count == PRODUCT Sale Count - How many Sold !
#
mdf <- data.frame(cty_name= c(ms_cty1,ms_cty2,ms_cty3,ms_cty4,ms_cty5,ms_cty6,ms_cty7,ms_cty8,ms_cty9,ms_cty10),
                  country_name= c(ms_cntry1,ms_cntry2,ms_cntry3,ms_cntry4,ms_cntry5,ms_cntry6,ms_cntry7,ms_cntry8,ms_cntry9,ms_cntry10),
                  p_sale_count= c(psale_1,psale_2,psale_3,psale_4,psale_5,psale_6,psale_7,psale_8,psale_9,psale_10),
                  p_sale_cost= c(pcost_1,pcost_2,pcost_3,pcost_4,pcost_5,pcost_6,pcost_7,pcost_8,pcost_9,pcost_10),
                  var_dates=c(dates_aa,dates_bb,dates_cc,dates_dd,dates_ee,dates_ff,dates_gg,dates_hh,dates_ii,dates_jj))
# #
head(mdf)
#
length(mdf$cty_name)
#
summary(mdf) # Summary of DF
#
str(mdf) # Structure of DF
# #

 
```
#Speeding up Code  

Efficiency Tradeoff --- 

Will we Multiply TWO Vectors 

OR 

Will we Multiply TWO DF Column Vectors 

There are ceratin text which recommend to Avoid "for Loops" 
or any other kind of iterations within R Code chunks  
At the same time the core dev team at R Studio recommends 
we need not avoid "for Loops" , thus its best to measure 
our own codes performance - specially if we want to use it again. 

We see below a brief intro to TIMING our code chunks...also
a brief intro to memory allocation.   

Further REFER -  
UCLA- https://stats.idre.ucla.edu/r/faq/how-can-i-time-my-code/  
Prof . Hadley Wickham - http://adv-r.had.co.nz/memory.html#object-size  
Also many other sources from the net.  
Rohit Dhankar claims no copyright to any of this code.   







```{r}

# Firstly lets create and multiply TWO Vectors 

p_sale_count<-c(psale_1,psale_2,psale_3,psale_4,psale_5,psale_6,psale_7,psale_8,psale_9,psale_10)

p_sale_cost<-c(pcost_1,pcost_2,pcost_3,pcost_4,pcost_5,pcost_6,pcost_7,pcost_8,pcost_9,pcost_10)


```



```{r}

# Start the clock!
ptm <- proc.time()

vec_gross_sale <- p_sale_count*p_sale_cost
summary(vec_gross_sale)

proc.time() - ptm 

#
# As seen below in our case 
# ELAPSED time - 1st 0.011 , 2nd - 0.012
# Thus the WALL CLOCK or REAL / ELAPSED 
# timings are almost same . 
#
# The USER TIME and SYSTEM TIME's in our case 
# add upto - 
# 1st - 0.008
# 2nd - 0.012

# Thus it would seem we are better off 
# with Vector Multiplication

# But we also need to consider 
# once we have the "vec_gross_sale" 
# we will need to add it to out "mdf"

# Kindly also note the Timings will
# differ for each system - also for each run 
# of the chunk of code on same sys

# Definition of user Time --- The ‘user time’ is the CPU time 
# charged for execution of user instructions of the calling process.
#
# REFER- https://stat.ethz.ch/R-manual/R-devel/library/base/html/proc.time.html



```


```{r}

# Now to multiply TWO Columns of the DF 
# Also called COLUMNAR VECTORS 

# Again start the clock!
ptm <- proc.time()

mdf$gross_sale<- mdf$p_sale_count*mdf$p_sale_cost

proc.time() - ptm 
#
str(mdf)
#
summary(mdf)
#
write.csv(mdf,file="Mkt_DATA_Files/mdf.csv") 
## Writes to Sub Directory - DATA_Files
#


```


```{r}


```




```{r}


```



```{r}


```



```{r}


```




```{r}


```




```{r}


```




```{r}


```




```{r}


```




```{r}


```




```{r}


```




```{r}


```




```{r}


```
